Local Spectral Clustering of Density Upper Level Sets

### 2019-05-23 neurips/local_spectral_density_clustering_neurips.tex
Tweak to discussion

\paragraph{Empirical behavior of PPR.} In Figure \ref{fig:fig2}, to drive home
the main implications of Theorems \ref{thm: misclassification_rate} and
\ref{thm: consistent_recovery_of_density_clusters}, we show the
behavior of PPR, normalized cut, and the density clustering algorithm of
\citet{chaudhuri2010} on the well known ``two moons'' dataset (with added 2d 
Gaussian noise), considered a prototypical success story for spectral clustering
algorithms. The first column shows the empirical density clusters $C[X]$
and $C'[X]'$ for a particular threshold $\lambda$ of the density function; the 
second column shows the cluster recovered by PPR; the third column shows the
global minimum normalized cut, computed according to the algorithm of
\citet{szlam2010}; and the last column shows a cut of the density cluster tree
estimator of \citet{chaudhuri2010}.  We can see the degrading ability of PPR to
recover density clusters as the two moons become less well-separated. Of 
particular interest is the fact that PPR fails to recover one of the moons even
when normalized cut still succeeds in doing so, supporting our claim from Remark
\ref{rmk: diameter}. Additionally, we note that the Chaudhuri-Dasgupta algorithm 
succeeds even when both PPR and normalized cut fail.  While our main message was
that PPR recovers geometrically well-conditioned density clusters, it would be  
interesting to establish that it \emph{only} recovers such clusters, a point we
return to below.

@@

There are an almost limitless number of ways to define what the ``right''
clustering is. In this paper, we have considered one such notion---density upper
level sets---and have detailed a set of natural geometric criteria which, when 
appropriately satisfied, translate to provable bounds on estimation of the
cluster by PPR. We do not, however, provide a theoretical lower bound showing
that our geometric conditions are required for successful recovery on an upper
level set. Although we investigate the matter empirically, this is a direction
for future work.

>>>>

\paragraph{Empirical behavior of PPR.} In Figure \ref{fig:fig2}, to drive home
the main implications of Theorems \ref{thm: misclassification_rate} and
\ref{thm: consistent_recovery_of_density_clusters}, we show the
behavior of PPR, normalized cut, and the density clustering algorithm of
\citet{chaudhuri2010} on the well known ``two moons'' dataset (with added 2d 
Gaussian noise), considered a prototypical success story for spectral clustering
algorithms. The first column shows the empirical density clusters $C[X]$
and $C'[X]'$ for a particular threshold $\lambda$ of the density function; the 
second column shows the cluster recovered by PPR; the third column shows the
global minimum normalized cut, computed according to the algorithm of
\citet{szlam2010}; and the last column shows a cut of the density cluster tree
estimator of \citet{chaudhuri2010}.  We can see the degrading ability of PPR to
recover density clusters as the two moons become less well-separated. Of 
particular interest is the fact that PPR fails to recover one of the moons even
when normalized cut still succeeds in doing so, supporting our claim from Remark
\ref{rmk: diameter}. Additionally, we note that the Chaudhuri-Dasgupta algorithm 
succeeds even when both PPR and normalized cut fail.  While our main message was
that PPR recovers geometrically well-conditioned density clusters, it would be  
interesting to establish that it \emph{only} recovers such clusters, a direction
for future work.

@@

There are an almost limitless number of ways to define what the ``right''
clustering is. In this paper, we have considered one such notion---density upper
level sets---and have detailed a set of natural geometric criteria which, when 
appropriately satisfied, translate to provable bounds on estimation of the
cluster by PPR. We do not, however, provide theory to show that our geometric
conditions are required for successful recovery of a density level set. This
seems to be supported by our empirical work, and it remains a direction for
future work.
